A Technique for Inter-Catalogue Recommendations

### 2014-06-11 paper2/paper.tex
Rework LatentConceptRecommender subsection, but not finished.

\subsection*{\em LatentConceptRecommender}
Although the $NeighborhoodRecommender$ alleviates some of the issues of 
content-based similarity, it is still primarily based on the descriptions of
products. Ideally, we would like to recommend products that are similar in more
fundamental ways. If a dress is elegant and modern, we would like to recommend 
another dress that is elegant and modern, even if those words are not explicitly
in the descriptions of either. These types of underlying concepts can be 
intangible and difficult to define. However, if we assume that the given
recommendations are truly based on such latent concepts, we can attempt to learn 
them by analyzing the graphs.

One common approach to learning latent semantic factors is to apply {\em
Singular Value Decomposition} (SVD) to a matrix of user data, as in many 
collaborative filtering methods [CITE]. This method requires a product-user
matrix, where each column represents the products viewed in a user session.
After SVD is applied to the matrix, the resulting three matrices are truncated
to the number of latent concepts desired. The resulting left matrix becomes a
product-concept matrix, where each row is a product represented in 
{\em concept-space}

In this experiment, there is no such product-user matrix to apply SVD.
However, since the recommendation graphs are constructed from user behavior,
we can use them to simulate user sessions. To do this, we use a random walk 
technique on the graph. We assume that the edges of a recommendation graph 
indeed connect products that a user would co-view. Thus traversing the edges 
of a graph is a reasonable simulation of a user session. For details on these
random walks, see the section titled Random Walks in the Implementation section
below.

With simulated user data for each recommendation graph, we can apply SVD on the
matrices from the two catalogues to learn the latent concepts for each product.
We want to find the products from one catalogue that are most similar in latent
concept-space to products in the other catalogue. However, since each user
session/random walk includes products exclusively from one of the catalogues, we
learn different latent spaces for each catalogue. Note that this would be an
issue even if we had the actual user data used to construct the two
recommendation graphs. The problem therefore becomes finding a mapping from one
concept-space to the other. If this can be accomplished, we can represent each
product of one catalogue in terms of concepts of the other catalogue and then
find the nearest neighbors of that catalogue by distance in concept-space.

To map concepts from one catalogue to another, we must revisit
content-similarity. In this case, we represent each concept (rather than each
product) as a vector of words. To represent a concept in terms of the
vocabulary, we calculate the value of each word by summing the
tf-idf of that word for each product $P$, weighted by the projection of $P$ on
our concept in concept-space. That is:

@@

>>>>

\subsection*{\em LatentConceptRecommender}
Although the $NeighborhoodRecommender$ leverages the information contained in
the recommendation graphs to some extent, it is still primarily based on product
descriptions. Presumably, the graphs are derivative of a collaborative filtering
scheme, and hence, reveal associations between products that are not evident
from the language of their descriptions. For example, a dress's description
might not capture if it is more conservative or more risque. Yet, this could be
the salient feature it has in common with an associated recommendation.  These
sort of abstract, perhaps ineffable, features can be essential to item-item
relations. We must therefore ask whether we can more fully utilize the given
graphical information to discover these latent features in order to choose
better inter-catalogue associations.

A common approach to learning latent features is to apply matrix factorization
techniques to user-item data, as in many CF methods \cite{Koren2009}. But,
user-item data, a prerequisite for such techniques, in not available to us here.
A key insight of the present work is that publicly available recommendation
graphs, such as those found on Macy's website, can be used to simulate user-item
data through the use of special random walks known as {\em Topic-Sensitive
PageRank}, a technique typically associated with web search
\cite{Haveliwala2002}.

The idea is that a user with a preference for a particular product is more
likely to also have a preference for one of its associated recommendations than
another arbitrary product. We can simulate this with a random walker that, by
construction has a preference for an item $A$, and therefore, begins her walk at
that item on the recommendation graph. At each step in the walk, she is equally
likely to visit any of the items that share an edge with the presently occupied
item or stay at the present item. There is also a small chance at each step that
the walker will teleport back to item $A$. In the language of Topic-Sensitive
PageRank, item $A$ is the sole nonzero entry in the walker's personalization
vector. We create our synthetic user-item data by performing such a random walk
for each node in the graph. That is, each node has a random walk in which it is
the sole nonzero entry in the walker's personalization vector. In this way, we
simulate a user for each item in each graph partition.

Constructing our random walk in this way provides an important benefit. In the
context of the random walk, the recommendation graph may be considered a {\em
discrete-time Markov chain}. By providing the walker a finite probability of
staying at each item, our chain becomes {\em aperiodic}. And since our walker
begins at $A$ and can only teleport to $A$, she is effectively confined to a
strongly connected subgraph of the recommendation graph. This subgraph is
therefore, by construction, {\em irreducible}.  Importantly, according to the
{\em Fundamental Theorem of Markov Chains}, any finite, aperiodic, irreducible
Markov chain has a unique stationary distribution \cite{Norris1998}.

We may then construct a synthetic $N \times N$ user-item matrix $M$ with rows
representing items and columns representing users, which are in fact the
stationary distributions discovered from our random walks. This matrix has at
least one notable advantage over most real user-item matrices---it is not
sparse.  Rather, each synthetic user has a finite value associated with every
item that is reachable from its origin item in the graph. Unlike the typical CF
scenario, we may therefore employ SVD as our matrix factorization technique
since our user-item matrix is virtually complete (i.e. nearly void of empty
entries).  Using SVD, we factor our user-item matrix as
\begin{align}
M = U \Sigma V^T,
\end{align}
where the rows of the $N \times N$ matrix $U$ and the $N \times N$ matrix $V$
are the items and users represented in an $N$-dimensional concept-space,
respectively, and the $N$ diagonal entries of $\Sigma$ are the singular values
associated with each concept.

Having transformed our items into concept-space, we now discard all but the $k$
leftmost columns of the matrix $U$. The idea is that these leftmost columns are
the most significant eigenvectors of concept-space and capture most, if not all,
of the relevant latent features. Meanwhile, the discarded columns are presumed
to contain mostly noise. This rank-reduction process yields matrices that are
typically written as
\begin{align}
M \approx U_k \Sigma_k V^T_k,
\end{align}
where the rows of $U_k$ and $V_k$ are the items and users represented in a
$k$-dimensional concept-space, respectively.  Rank-reducing following the
application of SVD in this way is a common technique used in information
retrieval and is often referred to as {\em Latent Semantec Indexing} (LSI)
\cite{Deerwester1990}.

Using random walks to simulate user data, followed by the application of LSI, is
indeed an effective way to encode information from a recommendation graph into a
latent feature space. But what is the relationship between latent feature
vectors derived from disconnected graph partitions? In principle, there is no
guarantee of any relationship between such latent spaces in the same way that
there is no guarantee of any relationship between two arbitrary, disconnected
recommendation graphs. Nonetheless, our hypothesis is that a relationship
between such latent spaces often will exist due to the effectiveness of the
underlying CF techniques that generate the graphs to capture meaningful
item-item associations. Put another way, we believe that the most fundamental
latent dimensions along which products in a category such as Women's swimsuits
vary will be common to many swimsuit catalogues. If this is indeed true, then it
might be possible to learn a mapping between these latent spaces.

If a mapping between the latent spaces associated with two catalogues were
known, we could translate all items into a common space and then choose
inter-catalogue edges between item pairs that are nearest to one another in
concept-space. To learn these mapping, we must revisit the notion of content
similarity. Once again, we utilize tf-idf vectors. This time, however, we weight
an item's tf-idf vector by the elements in the item's latent feature vector. In
this way, we construct a dense tf-idf vector for each concept for each graph
partition. Formally, we say that each concept $j$ is represented by 
\begin{align}
EXPRESSION FOR A CONCEPT'S TF-IDF VECTOR
\end{align}

In this case, we represent each concept (rather than each
product) as a vector of words. To represent a concept in terms of the
vocabulary, we calculate the value of each word by summing the tf-idf of that
word for each product $P$, weighted by the projection of $P$ on our concept in
concept-space. That is:

@@

\bibitem{Haveliwala2002}
    Taher Haveliwala.
    ``Topic-Sensitive PageRank.''
    \emph{Proceedings of the Eleventh International World Wide Web Conference}
(Honolulu, Hawaii) 2002.

\bibitem{Norris1998}
    James R. Norris.
    \emph{Markov Chains}. Cambridge University Press, 1998.

\bibitem{Deerwester1990}
    Scott Deerwester, Susan T. Dumais, George W. Furnas, Thomas K. Landauer,
Richard Harshman.
    ``Indexing by Latent Semantex Analysis.''
    \emph{American Society for Information Science} \textbf{41}, 391-407, 1990.
